---
title: "Accessing data from Google Docs"
author: "Nathan Brouwer"
date: "11/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this exercise is to make you familiar with how to download data from Google Sheets and to briefly review some key concepts R functions and coding concepts.

We'll do the following things

* download a list of RefSeq accessions from a Google sheet
* remove the NAs using na.omit()
* select out all but one isoform using duplicated()

## Packages 

```{r}
## Google sheets download package
# comment this out when you are done
# install.packages("googlesheets4")
library(googlesheets4)

# comp bio packages
library(seqinr)
library(rentrez)
library(compbio4all)
library(Biostrings)
```


## Download data

First, we need a web address (URL) for the spreadsheet with the data.

```{r}
spreadsheet_sp <- "https://docs.google.com/spreadsheets/d/1spC_ZA3_cVuvU3e_Jfcj2nEIfzp-vaP7SA5f-qwQ1pg/edit?usp=sharing" 
```

Second, we need to make sure we tall the package we aren't interested in checking user access credentials / authorization.

```{r}
# be sure to run this!
googlesheets4::gs4_deauth()   # <====== MUST RUN THIS
```

Third, we download our data.

#### NOTE!: sometimes Google Sheets or the function gets cranky and throws this error:

"Error in curl::curl_fetch_memory(url, handle = handle) : 
 Error in the HTTP2 framing layer"
 
If that happens, just re-run the code.

```{r}
# I include this again in case you missed is the first time : )
googlesheets4::gs4_deauth()  

# download
## NOTE: if you get an error, just run the code again
refseq_column <- read_sheet(ss = spreadsheet_sp, # the url
           sheet = "RefSeq_prot",                # the name of the worksheet
           range = "selenoprot!H1:H364",
           col_names = TRUE,
           na = "",                              # fill in empty spaces "" w/NA
           trim_ws = TRUE)
## NOTE: if you get an error, just run the code again

# for reasons we won't get into I'm going to do this
protein_refseq <- refseq_column$RefSeq_prot
```

Here's a snapshot of the results
```{r}
protein_refseq[1:10]
```
