---
title: "Line of best fit in R"
author: "Nathan Brouwer"
date: "11/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here's an example of how to make a graph with a line of best fit in R

Data are from Xiang et al. 2013. Identificaiton of Tibetan Specific Mutation in the Hypoxic gene EGLN1 and its contribvution to high-altitude adaptation.

Make the vectors
```{r}
population <- c("Nyingchi","Lhasa","Ch1amdo","Chamdo2","Xigaze","Shanna","Nagari","Nagqu")
altitude <- c(2750,3400,3500,3850,3890,3935,4700,4800)
allele_freq <- c(0.6992, 0.6429, 0.7234, 0.6549, 0.7151, 0.7119, 0.7583, 0.7542)
```


Make a dataframe

```{r}
tibetdf <- data.frame(pop = population,
                      alt = altitude,
                      freq = allele_freq)
```


Make a plot with plot()

```{r}
plot(freq ~ alt, data = tibetdf)
```

Calculate line of best with with lm()
```{r}
lobf <- lm(freq ~ alt, data = tibetdf)
```

Get slope and intercept for line
```{r}
lobf_slope_int <- coef(lobf)
```

Plot data with line
```{r}
plot(freq ~ alt, data = tibetdf, pch = 16, xlab = "Altitude (meters)", ylab = "C allele frequency %")
abline(lobf_slope_int, col =2, lwd = 2)
```

Calculate the correlation.  Compare to value in Table 1
```{r}
cor(tibetdf$freq, tibetdf$alt,method = "spearman")
```

