---
title: "Using R to investigate gene XXXX for sequence repeates"
author: "Nathan Brouwer"
date: "10/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Change the XXXXX of the title to your gene name.
1. Change the names and text appropriately to reflect your gene / protein.
1. Add the necessary code to make this script functional.
1. Download the PROTEIN sequence of your gene.
1. Adapting the code below, make 2 grids of 4 plots (8 plots total) exploring different settings for window size and the match threshold (nmatch)
1. Use the modified dotplot function to make a plot of your whole gene with only one part shown.
1. Use the modified dotplot function to plot just a focal section of your gene.
1. Make sure that all plots have titles relevant to your gene.

This script is based on the Shroom example used in a previous Portfolio assignment.




## Preliminaries

### Load packages

```{r}
library(seqinr)
library(rentrez)
library(compbio4all)
library(Biostrings)
```



### Get gene XXXXX data

Download sequence of NP_000363.
```{r}
tyr_fasta <- rentrez::entrez_fetch(id = "NP_000363",
                                      db = "protein", 
                                      rettype="fasta")
```

Clean and set up sequence as vector.

```{r}
tyr_vector <- fasta_cleaner(tyr_fasta)
```

Confirm set up.
```{r}
str(tyr_vector)
```

Make a 2 x 2 grid of dotplots to explore effect of changing window size and nmatch
```{r}
# set up 2 x 2 grid, make margins thinner
par(mfrow = c(2,2), 
    mar = c(0,0,2,1))

# plot 1:  - Defaults
dotPlot(tyr_vector, tyr_vector, 
        wsize = 1, 
        nmatch = 1, 
        main = "tyr Defaults")

# plot - size = X, nmatch = X
dotPlot(tyr_vector, tyr_vector, 
        wsize = 10, 
        nmatch = 1, 
        main = " tyr- size = 10, nmatch = 1")

# plot 3:  - size = , nmatch = 
dotPlot(tyr_vector,tyr_vector, 
        wsize = 10, 
        nmatch = 5, 
        main = "tyr - size = 10, nmatch = 5")

# plot 4: size = , nmatch = 
dotPlot(tyr_vector, tyr_vector, 
        wsize = 20,
        nmatch = 5,
        main = "tyr - size = 20, nmatch = 5")

# reset par() - run this or other plots will be small!
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))
```


Make ANOTER 2 x 2 grid of dotplots to explore effect of changing window size and nmatch
```{r}
# set up 2 x 2 grid, make margins thinner
par(mfrow = c(2,2), 
    mar = c(0,0,2,1))

# plot 1: 
dotPlot(tyr_vector, tyr_vector, 
        wsize = 30, 
        nmatch = 5, 
        main = "tyr 30 / 5")

# plot 2 tyr - size = 30, nmatch =10
dotPlot(tyr_vector,tyr_vector, 
        wsize = 30, 
        nmatch = 5, 
        main = "tyr - size = 30, nmatch = 10")

# plot 3: tyr - size = 5, nmatch = 2
dotPlot(tyr_vector, tyr_vector, 
        wsize = 5, 
        nmatch = 2, 
        main = "tyr - size = 5, nmatch = 2")

# plot 4: size = 12, nmatch = 4
dotPlot(tyr_vector, tyr_vector, 
        wsize = 12,
        nmatch = 4,
        main = "tyr - size = 12 , nmatch = 4")

# reset par() - run this or other plots will be small!
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))
```



##  Best plot using normal dotplot

This is the most interesting gene abcdefghijklmnop dotplot.
```{r}
# be sure to run par - re-run just in case
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

dotPlot(tyr_vector, 
        tyr_vector,
        wsize = 30, 
        nmatch = 5, )
```



## New dotplot function

The code below makes a new dotplot function

Make new function 
```{r}
dot_plot <- function(seq1, seq2, wsize = 1, wstep = 1, nmatch = 1, col = c("white", 
    "black"), xlab = deparse(substitute(seq1)), ylab = deparse(substitute(seq2)), 
    ...) {
  
  # make sure input works
    if (nchar(seq1[1]) > 1) 
        stop("seq1 should be provided as a vector of single chars")
    if (nchar(seq2[1]) > 1) 
        stop("seq2 should be provided as a vector of single chars")
    if (wsize < 1) 
        stop("non allowed value for wsize")
    if (wstep < 1) 
        stop("non allowed value for wstep")
    if (nmatch < 1) 
        stop("non allowed value for nmatch")
    if (nmatch > wsize) 
        stop("nmatch > wsize is not allowed")
  
  # internal function
    mkwin <- function(seq, wsize, wstep) {
        sapply(seq(from = 1, to = length(seq) - wsize + 1, by = wstep), 
            function(i) c2s(seq[i:(i + wsize - 1)]))
    }
    wseq1 <- mkwin(seq1, wsize, wstep)
    wseq2 <- mkwin(seq2, wsize, wstep)
    if (nmatch == wsize) {
        xy <- outer(wseq1, wseq2, "==")
    }
    else {
        "%==%" <- function(x, y) colSums(sapply(x, s2c) == sapply(y, 
            s2c)) >= nmatch
        xy <- outer(wseq1, wseq2, "%==%")
    }
    
    # compile output in list
    out <- list(x = seq(from = 1, to = length(seq1), length = length(wseq1)),
                y = seq(from = 1, to = length(seq2), length = length(wseq2)),
                z = xy)
}
```

## Full-length plot

Use new function on the full-length protein, save output (doesn't autoplot)
```{r}
my_dot_out <- dot_plot(tyr_vector,
                       tyr_vector, 
        wsize = 30, 
        wstep = 1,
        nmatch = 5)
```

Get rid of upper triangular portion
```{r}
# don't change this
my_dot_out$z[lower.tri(my_dot_out$z)] <- FALSE

```

Do some weird prep (don't worry about it)
```{r}
# don't change this
my_dot_out$z <- my_dot_out$z[, nrow(my_dot_out$z):1]
```

Plot using image() command
```{r}
# seriously - it will drive you crazy if you forget about this
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

# plot with image()
image(x = my_dot_out$x, 
      y = my_dot_out$y, 
      z = my_dot_out$z)
```

## Focal subplot

Use new function on the full-length protein, save output (doesn't autoplot)
```{r}
my_dot_out <- dot_plot(tyr_vector,
                       tyr_vector, 
        wsize = 30, 
        wstep = 1,
        nmatch = 5)
```

Get rid of upper triangular portion
```{r}
# don't change this
my_dot_out$z[lower.tri(my_dot_out$z)] <- FALSE

```

Do some weird prep (don't worry about it)
```{r}
# don't change this
my_dot_out$z <- my_dot_out$z[, nrow(my_dot_out$z):1]
```

Plot using image() command
```{r}
# seriously - it will drive you crazy if you forget about this
par(mfrow = c(1,1), 
    mar = c(4,4,4,4))

# plot with image()
# don't change this
image(x = my_dot_out$x, 
      y = my_dot_out$y, 
      z = my_dot_out$z)
```
